---
title: "Human Activity Recognition"
author: "Andrei Kuebar"
date: "Sunday, April 17, 2016"
output: html_document
---

# Summary
In this project we are dealing with the human activity data gathered by means of commonly available fitness trackers such as Jawbone or Fitbit.

More details on the data are here: [link](http://groupware.les.inf.puc-rio.br/har)

Model we are building is going to predict the fashion in which the exercise was performed. There are total 5 possiblities, out of which 4 are incorrect technique and 1 correct. They are marked "A", "B", "C", "D" and "E". "A" being the correct one and others incorrect.

# Get the data
Download the training and test datasets from:

[train data](https://eventing.coursera.org/api/redirectStrict/BShbnUGy5UftRGyQm7vg-mQrT2Wtz7-GeN_bEl6rLLxEoNx_6xxFkqldMUelho0FQOK8ibQ3b-zHbXfsQ90yvg.Du4b6W90rCwPGeAl54V8zg.7789zd8s0B5I-0DHZl-FkYAqeHkJqfgv58FSNVpggClDZmRp6btrp9inLWtO4dYmZrLCAnXGW-BXypZx75giSBZd-teT13Ywr9BwTk8DL7Dhz_Lmz3LtYEALyDDDsDlPxfgROk9NrNQ8qqtEUpdRI1vF3xQ2Snor2ZD97gXx9V7d6vsBn3CZF7hYIvT-ky-8fJFmzh_lfPXAo0uIeJeHHJbXpg8WcCiWjfeLfp8VT1kUcB-9HxvxrX8ntQe6CPIz0Ag0jsYlVHN8LS_x3k7-l7HMUZG8AIOrGeF6PuisRuHfn7QgBKooFjiMcKcTeGFSIJl2SzZBKoDMlzDPq292QtIVKbIH8MhKDcZjrU0Kkd1QlaY0R2rTvn3zbBXVJzOmziRRRQA9KuhH5Y5x54zCXA)

[test data](https://eventing.coursera.org/api/redirectStrict/BShbnUGy5UftRGyQm7vg-mQrT2Wtz7-GeN_bEl6rLLxEoNx_6xxFkqldMUelho0FQOK8ibQ3b-zHbXfsQ90yvg.Du4b6W90rCwPGeAl54V8zg.7789zd8s0B5I-0DHZl-FkYAqeHkJqfgv58FSNVpggClDZmRp6btrp9inLWtO4dYmZrLCAnXGW-BXypZx75giSBZd-teT13Ywr9BwTk8DL7Dhz_Lmz3LtYEALyDDDsDlPxfgROk9NrNQ8qqtEUpdRI1vF3xQ2Snor2ZD97gXx9V7d6vsBn3CZF7hYIvT-ky-8fJFmzh_lfPXAo0uIeJeHHJbXpg8WcCiWjfeLfp8VT1kUcB-9HxvxrX8ntQe6CPIz0Ag0jsYlVHN8LS_x3k7-l7HMUZG8AIOrGeF6PuisRuHfn7QgBKooFjiMcKcTeGFSIJl2SzZBKoDMlzDPq292QtIVKbIH8MhKDcZjrU0Kkd1QlaY0R2rTvn3zbBXVJzOmziRRRQA9KuhH5Y5x54zCXA)

```{r}
training <- read.csv("pml-training.csv")
testing <- read.csv("pml-testing.csv")
```

# Clean the data
### Remove all column containing NA or empty strings
```{r}
trColNA<-sapply(training, function (x) ( sum(is.na(x))>0|sum(x=="")>0 ) )
training_NNA<-training[,names(training)[!trColNA]]
dim(training_NNA)
```

### Pick only the variables which contain measurements
```{r}
pred<-names(training_NNA)[grep("roll|pitch|yaw|gyros|^accel|^magnet",names(training_NNA))]
names(training_NNA[,pred])
```

# Fit the model

### Enable parallelization
```{r eval=FALSE}
library(parallel)
library(doParallel)
cluster <- makeCluster(detectCores() - 1)
registerDoParallel(cluster)
#stopCluster(cluster)
```

### Configure the trainControl object to define cross-validation method.
```{r eval=FALSE}
fitControl <- trainControl(method = "boot",
                           number = 10,
                           allowParallel = TRUE)
```

### After trying out several different models, RandomForest model proved to have the highest accuracy.
```{r cache=TRUE, eval=FALSE}
modFit<-train(training_NNA[,pred], training_NNA$classe,  method="rf", trControl = fitControl)
modFit
```
```{r eval=FALSE}
  mtry  Accuracy   Kappa      Accuracy SD   Kappa SD   
   2    0.9936879  0.9920126  0.0013031752  0.001645974
  25    0.9936011  0.9919028  0.0009676512  0.001222974
  48    0.9873407  0.9839790  0.0027990771  0.003543746
```

### We can now generate the predictions on the test data
```{r eval=FALSE}
colnames<-as.character(1:20)
predTest<-as.character(predict(modFit,testing[,pred]))
names(predTest)<-colnames
predTest
```

```{r eval=FALSE}
  1   2   3   4   5   6   7   8   9  10  11  12  13  14  15  16  17  18  19  20 
"B" "A" "B" "A" "A" "E" "D" "B" "A" "A" "B" "C" "B" "A" "E" "E" "A" "B" "B" "B" 
```










